### 2023-09-01 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@ac86af753d6ea60ef333d44f62aa0dd3fb7860c4 🚀

Bats, the second most diverse groups of mammals, are found worldwide;
gradients in their species richness generally track broader patterns of
mammal diversity,\textsuperscript{\textbf{Tanalgo2022Mapping?}} with a
striking Neotropical hotspot (especially in the Amazon basin) and a
secondary hotspot centered in Indochina. These hotspots of bat diversity
are generally presumed to be hotspots of viral adaptive radiation, and
therefore areas of concern for human
health.\textsuperscript{\textbf{Anthony2017GloPat?},\textbf{Olival2017HosVir?}}
However, the hotspots of known bat betacoronavirus hosts show a distinct
pattern, with primary hotspots (both in terms of area and higher values)
of host richness situated in southeast Asia, parts of southern Europe,
and to a lesser extent parts of Africa in the -25-0 range of latitudes
(fig.~\ref{fig:richness}; top). Although hundreds of species likely host
undiscovered betacoronaviruses, machine learning predictions have
suggested that these undiscovered reservoirs should follow the same
diversity gradient.\textsuperscript{\textbf{Becker2022OptPre?}} In
principle, these hotspots of locally-diverse, virus-rich bat communities
should drive more adaptive diversification in their viruses.

@@

Bats---and the spillover of their viruses---are also sensitive to
anthropogenic factors others than climate change, including
deforestation and other kinds of habitat loss, increased stress, and
greater contact with potential bridge hosts like domesticated
species.\textsuperscript{\textbf{Alves2018GeoVar?},\textbf{Treitler2016EffLoc?},\textbf{Rulli2021LanCha?},\textbf{Mendenhall2014PreBio?}}
This represents a challenge for both conservation strategies and
pandemic prevention,\textsuperscript{\textbf{Amman2011InvRol?}} but
identifying areas at risk, and protecting the health of bats and
ecosystems within those zones, can be a win-win intervention for
both.\textsuperscript{\textbf{Hopkins2021HowIde?},\textbf{Plowright2021LanUse?},\textbf{OHHLEP2022OneHea?}}
As we scale these predictions down in space to finer spatial resolutions
to guide public health actions,\textsuperscript{18} the incorporation of
human activity predictors will become more
importyant.\textsuperscript{19}

@@

For all bat hosts of \emph{Betacoronavirus}, we extracted their
predicted viral sharing network, generated from a previously published
generalized additive mixed model of virus sharing by a tensor function
of phylogenetic distance and geographic range overlap across
mammals.\textsuperscript{\textbf{Albery2020PreGlo?}} This network stores
pairwise values of viral community similarity, measured for all hosts
(to maintain consistency with teh phylogenetic diversity measure) across
all viruses; therefore, we consider that it accounts for some overall
similarity in the way hosts deal with viruses, and not only
betacoronaviruses. There is empirical evidence that capacity for
cross-species transmission even between divergent species is generally
high,\textsuperscript{20} especially for
beta-coronaviruses.\textsuperscript{13} To project viral sharing values
into a single value for every pixel, we averaged the pairwise scores.
High values of the average sharing propensity means that this specific
extant bat assemblage is likely to be proficient at exchanging viruses.

@@

\leavevmode\hypertarget{ref-Muylaert2022Present}{}%
\CSLLeftMargin{18. }
\CSLRightInline{Muylaert, R. L. \emph{et al.} Present and future
distribution of bat hosts of sarbecoviruses: Implications for
conservation and public health. \emph{Proceedings of the Royal Society
B: Biological Sciences} \textbf{289}, 20220397 (2022).}

\leavevmode\hypertarget{ref-Ka-WaiHui2006Reasons}{}%
\CSLLeftMargin{19. }
\CSLRightInline{Ka-Wai Hui, E. Reasons for the increase in emerging and
re-emerging viral infectious diseases. \emph{Microbes and Infection}
\textbf{8}, 905--916 (2006).}

\leavevmode\hypertarget{ref-Mollentze2020Viral}{}%
\CSLLeftMargin{20. }
\CSLRightInline{Mollentze, N. \& Streicker, D. G. Viral zoonotic risk is
homogenous among taxonomic orders of mammalian and avian reservoir
hosts. \emph{Proceedings of the National Academy of Sciences}
\textbf{117}, 9423--9430 (2020).}

>>>>

Bats, the second most diverse groups of mammals, are found worldwide;
gradients in their species richness generally track broader patterns of
mammal diversity,\textsuperscript{18} with a striking Neotropical
hotspot (especially in the Amazon basin) and a secondary hotspot
centered in Indochina. These hotspots of bat diversity are generally
presumed to be hotspots of viral adaptive radiation, and therefore areas
of concern for human health.\textsuperscript{11,19} However, the
hotspots of known bat betacoronavirus hosts show a distinct pattern,
with primary hotspots (both in terms of area and higher values) of host
richness situated in southeast Asia, parts of southern Europe, and to a
lesser extent parts of Africa in the -25-0 range of latitudes
(fig.~\ref{fig:richness}; top). Although hundreds of species likely host
undiscovered betacoronaviruses, machine learning predictions have
suggested that these undiscovered reservoirs should follow the same
diversity gradient.\textsuperscript{20} In principle, these hotspots of
locally-diverse, virus-rich bat communities should drive more adaptive
diversification in their viruses.

@@

Bats---and the spillover of their viruses---are also sensitive to
anthropogenic factors others than climate change, including
deforestation and other kinds of habitat loss, increased stress, and
greater contact with potential bridge hosts like domesticated
species.\textsuperscript{\textbf{Alves2018GeoVar?},\textbf{Treitler2016EffLoc?},\textbf{Rulli2021LanCha?},\textbf{Mendenhall2014PreBio?}}
This represents a challenge for both conservation strategies and
pandemic prevention,\textsuperscript{\textbf{Amman2011InvRol?}} but
identifying areas at risk, and protecting the health of bats and
ecosystems within those zones, can be a win-win intervention for
both.\textsuperscript{\textbf{Hopkins2021HowIde?},\textbf{Plowright2021LanUse?},\textbf{OHHLEP2022OneHea?}}
As we scale these predictions down in space to finer spatial resolutions
to guide public health actions,\textsuperscript{21} the incorporation of
human activity predictors will become more
importyant.\textsuperscript{22}

@@

For all bat hosts of \emph{Betacoronavirus}, we extracted their
predicted viral sharing network, generated from a previously published
generalized additive mixed model of virus sharing by a tensor function
of phylogenetic distance and geographic range overlap across
mammals.\textsuperscript{\textbf{Albery2020PreGlo?}} This network stores
pairwise values of viral community similarity, measured for all hosts
(to maintain consistency with teh phylogenetic diversity measure) across
all viruses; therefore, we consider that it accounts for some overall
similarity in the way hosts deal with viruses, and not only
betacoronaviruses. There is empirical evidence that capacity for
cross-species transmission even between divergent species is generally
high,\textsuperscript{23} especially for
beta-coronaviruses.\textsuperscript{13} To project viral sharing values
into a single value for every pixel, we averaged the pairwise scores.
High values of the average sharing propensity means that this specific
extant bat assemblage is likely to be proficient at exchanging viruses.

@@

\leavevmode\hypertarget{ref-Tanalgo2022Mapping}{}%
\CSLLeftMargin{18. }
\CSLRightInline{Tanalgo, K. C., Oliveira, H. F. M. \& Hughes, A. C.
Mapping global conservation priorities and habitat vulnerabilities for
cave-dwelling bats in a changing world. \emph{Science of The Total
Environment} \textbf{843}, 156909 (2022).}

\leavevmode\hypertarget{ref-Olival2017Host}{}%
\CSLLeftMargin{19. }
\CSLRightInline{Olival, K. J. \emph{et al.} Host and viral traits
predict zoonotic spillover from mammals. \emph{Nature} \textbf{546},
646--650 (2017).}

\leavevmode\hypertarget{ref-Becker2022Optimising}{}%
\CSLLeftMargin{20. }
\CSLRightInline{Becker, D. J. \emph{et al.} Optimising predictive models
to prioritise viral discovery in zoonotic reservoirs. \emph{The Lancet
Microbe} (2022)
doi:\href{https://doi.org/10.1016/S2666-5247(21)00245-7}{10.1016/S2666-5247(21)00245-7}.}

\leavevmode\hypertarget{ref-Muylaert2022Present}{}%
\CSLLeftMargin{21. }
\CSLRightInline{Muylaert, R. L. \emph{et al.} Present and future
distribution of bat hosts of sarbecoviruses: Implications for
conservation and public health. \emph{Proceedings of the Royal Society
B: Biological Sciences} \textbf{289}, 20220397 (2022).}

\leavevmode\hypertarget{ref-Ka-WaiHui2006Reasons}{}%
\CSLLeftMargin{22. }
\CSLRightInline{Ka-Wai Hui, E. Reasons for the increase in emerging and
re-emerging viral infectious diseases. \emph{Microbes and Infection}
\textbf{8}, 905--916 (2006).}

\leavevmode\hypertarget{ref-Mollentze2020Viral}{}%
\CSLLeftMargin{23. }
\CSLRightInline{Mollentze, N. \& Streicker, D. G. Viral zoonotic risk is
homogenous among taxonomic orders of mammalian and avian reservoir
hosts. \emph{Proceedings of the National Academy of Sciences}
\textbf{117}, 9423--9430 (2020).}
